ChemKED: a human- and machine-readable data standard for chemical kinetics experiments

### 2017-10-08 rebuttal.tex

\begin{response}
    %
    We removed the extraneous comma on \textbf{page 8}.
    %
\end{response}

\subsection*{PyKED architecture}

\begin{reviewer}
    %
    16.~The Python package Cerberus is used to validate the format and content of the given ChemKED
    file. How does this package handles/reports the inconsistencies with the schemas? As far as my
    local test went with giving a corrupted YAML file to the ChemKED initiator, the package just
    spits out a Python error, which might be not the most user-friendly solution.
    %
\end{reviewer}

\begin{response}
    %
    We recognize that this is a limitation of the current version of PyKED. The technical reason for
    this in the Cerberus package has very recently been resolved (see
    \url{https://github.com/pyeve/cerberus/issues/278}), so we plan to update the error-reporting
    mechanism in PyKED to report well-formatted errors that have a detailed description of the
    error. We have opened a public issue to track the progress on this:
    \url{https://github.com/pr-omethe-us/PyKED/issues/78}. Thank you for suggesting this
    improvement.
    %
\end{response}

@@

    Finally, each released version of PyKED is permanently archived on Zenodo
    (\url{https://doi.org/10.5281/zenodo.597935}) so that it is accessible should a user wish to use
    an old version to solve a compatibility issue temporarily.
    %
\end{response}

\begin{reviewer}
    %
    18.~Page 7, line 12, \texttt{mole-percent}, \texttt{mole-fraction}, and \texttt{mass-fraction}
    include a dash in their name, whereas this is not the case for the example on page 5 line 48 and
    in the description of the composition element on page 4, line 47.

    19.~Same issue as \#18 holds for page 7, line 17.
    %
\end{reviewer}

\begin{response}
    %
    This was a typo in the manuscript. All of the keywords here include spaces in their name, both
    in the YAML schema and in the PyKED interface. This has been corrected in the manuscript on
    \textbf{page 9} and we have added a test to the PyKED package to ensure the values are correct.
    %
\end{response}

### 2017-10-08 rebuttal.tex

\begin{response}
    %
    This was a typo in the manuscript. All of the keywords here include spaces in their name, both
    in the YAML schema and in the PyKED interface. This has been corrected in the manuscript on
    \textbf{page 9} and we have added a test to the PyKED package to ensure the values are correct.
    %
\end{response}

\begin{reviewer}
    %
    20.~I think it would be more clear if \texttt{volume\_history}, \texttt{compression\_time},
    \texttt{compressed\_temperature}, and \texttt{compressed\_pressure} would be combined.

    The line: `if the ChemKED file encodes an RCM experiment' is unnecessarily repeated.
    %
\end{reviewer}

\begin{response}
    %
    We have consolidated the RCM-specific information into separate list for ChemKED on
    \textbf{page 7--8} and PyKED on \textbf{page 9}, which eliminated the
    duplicated phrase.
    %
\end{response}

@@

\begin{reviewer}
    %
    21.~Page 7, line 39: ``Future versions of PyKED may support this feature via, e.g. online
    lookup''. There are two main caveats with online lookups: a) It will make the code slower if the
    internet connection is not strong, b) an internet connection is necessary for the code to be
    executed.
    %
\end{reviewer}

\begin{response}
    %
    In general, both of these points are correct. However, we believe there are several mitigating
    factors that make the online lookups worthwhile. First, we explicitly test that our code fails
    gracefully in the absence of an internet connection. No internet connection is necessary to run
    the code at this time, and aside from specialized features, we do not anticipate that an
    internet connection will ever be required to run the basic software.

    Second, the specific feature mentioned here will be something that the user explicitly requests (e.g., by
    attempting to convert mass to mole fractions). Thus, we expect that they are willing to wait for
    the connection to complete to use the feature. Third, we expect that the amount of data
    transmitted for such a request would be small, on the order of a few kB, so that even slow
    internet connections would be able to handle such requests. Finally, in addition to
    internet-based options, we are also investigating the use of standalone packages, such as RDKit,
    that would enable local retrieval of such information.
    %
\end{response}

### 2017-10-08 rebuttal.tex

    Users can create Excel or CSV files containing all or some of the datapoints in the ChemKED
    instance by using the \verb|get_dataframe()| function, and the associated functionality in
    Pandas (\verb|DataFrame.to_csv()| and \verb|DataFrame.to_excel()|). Since this functionality
    already exists, we do not wish to reinvent the wheel by writing our own functions that would
    have to be maintained and tested. Finally, we would note that data can also be plotted and
    viewed directly from Python using existing functionality in the Pandas DataFrame class
    (\verb|DataFrame.plot()|).
    %
\end{response}

\subsection*{Usage examples}
\subsubsection*{RCM modelling with varying reactor volume}

\begin{reviewer}
    %
    23.~Page 10, line 29--30: ``A user might want to load the information'' Avoid the term `might'.
    It's one of the main goals of using ChemKED and its associated Python-package PyKED to load the
    information necessary for simulating the experiment.
    %
\end{reviewer}

\begin{response}
    %
    We have modified the text on \textbf{page 12} to describe this example as an application.
    %
\end{response}

@@

\begin{reviewer}
    %
    24.~The Python code snippets are not always initiated in the text.
    %
\end{reviewer}

\begin{response}
    %
    We have corrected this oversight, with more explanation and introduction of the Python
    examples on \textbf{pages 12--17}.
    %
\end{response}

@@

\begin{reviewer}
    %
    25.~Page 10, line 55: add more commentary text. It is not clear that those lines create the
    necessary instances for Cantera, rather than ChemKED.
    %
\end{reviewer}

\begin{response}
    %
    We have expanded the explanation for all of the Python examples to give more detail of the
    process of creating the simulation.
    %
\end{response}

\begin{reviewer}
    %
    26.~Page 11, line 7: Where is this `air.xml' file coming from. This is not clear from the text.
    %
\end{reviewer}

\begin{response}
    %
    \verb|air.xml| is an input file included with Cantera, representing an average composition of
    air. We have modified the text on \textbf{page 13} to include this information.
    %
\end{response}

### 2017-10-08 rebuttal.tex

\begin{response}
    %
    \verb|air.xml| is an input file included with Cantera, representing an average composition of
    air. We have modified the text on \textbf{page 13} to include this information.
    %
\end{response}

\subsubsection*{Shock tube modelling with constant volume}

\begin{reviewer}
    %
    27.~The table overlapping page 11 and 12 can be reduced by truncating the \texttt{DataPoints},
    similarly as is done with the table starting at page 9, line 31 and ending at page 10, line 27.
    %
\end{reviewer}

\begin{response}
    %
    We have truncated the datapoints list on \textbf{pages 12 and 16} as the reviewer suggests.
    %
\end{response}

@@

\begin{reviewer}
    %
    28. Python code snippet on page 13: (a.) Add more commentary text. (b.) Changing the example to
    enable multithreading might make it more appealing to start using ChemKED as a data format in
    addition with Cantera as a simulation tool.
    %
\end{reviewer}

\begin{response}
    %
    We have modified the example on \textbf{pages 16--17} to include more
    explanatory text and to use Python's built-in
    multiprocessing library to run the simulations on many processors. We note that if Cantera has
    been built with multi-threaded BLAS and LAPACK libraries (such as the Intel MKL), then the 0-D
    solver will automatically utilize the multithreading capabilities of those libraries. Therefore,
    using the multiprocessing library may not give much benefit in terms of overall speedup,
    depending on the complexity of the chemical kinetic model (i.e., how long the simulation
    takes to run), how many processors are available, and the availability of the multi-threaded
    BLAS and LAPACk libraries. However, we have not explained this detail of Cantera in the text to
    avoid confusing the reader.
    %
\end{response}

### 2017-10-08 pyked.tex
mechanism -> kinetic model
adds page ref and number of processes for #43

Then, we define a function that will be mapped onto each job. This function takes the initial
temperature, pressure, and mole fractions as input and returns the simulated ignition delay time,
defined as the time the temperature increases by \SI{400}{\K} over the initial temperature, a
simplified definition for this example. In general, the user could process the mole fraction of
\ce{OH^*} (provided that the mechanism included accurate chemistry for \ce{OH^*}) to match the
definition of ignition delay in the experiments. As before, the \verb|tbuoh-mechanism.cti| file must
be supplied by the user:
%
\begin{pythonbox}
def run_simulation(T, P, X):
    gas = ct.Solution('tbuoh-mechanism.cti')
    gas.TPX = T, P, X
    reac = ct.IdealGasReactor(gas)
    netw = ct.ReactorNet([reac])
    while reac.T < T + 400:
        netw.step()

@@

Finally, we create the processor \pybox|Pool| and send the jobs out to run:
%
\begin{pythonbox}
with Pool(12) as pool:
    ignition_delays = pool.starmap(run_simulation, initial_conditions)

>>>>

Then, we define a function that will be mapped onto each job. This function takes the initial
temperature, pressure, and mole fractions as input and returns the simulated ignition delay time,
defined as the time the temperature increases by \SI{400}{\K} over the initial temperature, a
simplified definition for this example. In general, the user could process the mole fraction of
\ce{OH^*} (provided that the kinetic model includes accurate chemistry for \ce{OH^*}) to match the
definition of ignition delay in the experiments. As before, the \verb|tbuoh-mechanism.cti| file must
be supplied by the user:
%
\begin{pythonbox}
def run_simulation(T, P, X):
    gas = ct.Solution('tbuoh-mechanism.cti')
    gas.TPX = T, P, X
    reac = ct.IdealGasReactor(gas)
    netw = ct.ReactorNet([reac])
    while reac.T < T + 400:
        netw.step()

@@

Finally, we create the processor \pybox|Pool| (with 12 processes) and send the jobs out to run:
%
\begin{pythonbox}
with Pool(12) as pool:
    ignition_delays = pool.starmap(run_simulation, initial_conditions)
